---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# censusviz

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/censusviz)](https://CRAN.R-project.org/package=censusviz)
<!-- badges: end -->

The goal of censusviz is to:

(1) converting historical demographic data from IPUMS into an accessible format for analysis and joining it with geospatial census tract data so that it is easy to visualize on a map, 

(2) visualize the distribution of demographic data by census tract for a specified region and year on a leaflet map, 

(3) create a simple shiny app that allows users to interactively explore changes over time in demographics through leaflet maps and data tables. 

## Installation

`censusviz` is hosted on GitHub and can be installed by running the following function:

``` {r, message = FALSE}
remotes::install_github("rporta23/censusviz")
```

```{r example}
library(censusviz)
```

## Contributors

- [Irene Foster](https://github.com/i-m-foster)
- [Catherine Park](https://github.com/CJParkNW)
- [Rose Porta](https://github.com/rporta23)
